Estimating changes in temperature distributions in a large ensemble of climate simulations using quantile regression

### 2018-05-24 paper_revision.tex
Added Ryans comments

\affiliation{University of Chicago, Chicago, USA} 
% \affiliation{University of Illinois at Urbana-Champain, Urbana, USA} 
\email{mahaugen@uchicago.edu}
\extraauthor{Ryan L. Sriver}
    \extraaffil{University of Illinois at Urbana-Champain, Urbana, USA}
% \abstract{
% Understanding future changes in extreme temperature events in a transient climate is inherently challenging. Because a single model simulation is generally insufficient to empirically characterize the statistical properties of underlying processes, many recent efforts have generated ensembles of repeated simulations (with different initial conditions providing independence) to alleviate that data scarcity. We present here a new method for making use of these ensembles that allows characterizing changes in temperature variability for each given time with high confidence by analyzing all temporal observations simultaneously, assuming that temperature distributions vary smoothly both day to day over an annual cycle and over longer secular trends. To demonstrate our method's utility, we analyze an ensemble of 50 simulations of the Community Earth System Model (CESM) under a scenario of increasing radiative forcing to 2100, focusing on North America in particular. The results both verify aspects of climate system behavior known from previous studies, including that wintertime daily temperature variance generally decreases in the continental mid- and high- latitudes ($> 45^{\circ}$), and also elucidate new features. For example, we find that wintertime continental temperature distributions change in a spatial dipole pattern, becoming more left-skewed in the north ($> 45^{\circ}$) and more symmetric at lower latitudes; while changes in summertime distributions show no clear spatial pattern. That is, winter cold extremes become more variable relative to the overall distributions. Although the examples above concern temperature only, the technique is sufficiently general that it can be used to generate precise estimates of distribution changes in a broad range of climate variables by exploiting the power of ensembles.}

\abstract{
Understanding future changes in extreme temperature events in a transient climate is inherently challenging. A single model simulation is generally insufficient to empirically characterize the statistical properties of the underlying physical processes governing the climate. Ensembles of repeated simulations (with different initial conditions providing independence) greatly expand the amount of data available, which in turn allows new approaches for characterizing changes in extremes. We present here one such new approach, using ensembles that allow characterizing changes in temperature distributions using a continuous representation of seasonality rather than breaking the dataset into seasonal blocks. That is, we assume that temperature distributions evolve smoothly both day-to-day over an annual cycle and year-to-year over longer secular trends. To demonstrate our method's utility, we analyze an ensemble of 50 simulations of the Community Earth System Model (CESM) under a scenario of increasing radiative forcing to 2100, focusing on North America. The results both confirm aspects of climate system behavior known from previous studies and also elucidate new features. Confirming results include that daily temperature bulk variability generally decreases in wintertime in the continental mid- and high-latitudes ($> 40^{\circ}$). One new result is that these same wintertime distributions, the low tails ``stick'', i.e.\ experience a lesser reduction in variability, producing a more negative skew. Although the examples above concern temperature only, the technique is sufficiently general that it can be used to generate precise estimates of distribution changes in a broad range of climate variables by exploiting the power of ensembles.}

@@

Climate time series have generally been assumed to be separable into two components: randomness inherent in the underlying physical processes, which we call natural variability, and climatic trends, e.g. in the form of forced secular trends that follow from increasing concentrations of greenhouse gases or seasonal trends.
Recently, the degree to which natural variability may itself be changing has received significant scientific interest \citep[e.g.][]{trenberth2011attribution, donat2012shifting,deser2012communication, thompson2015quantifying, kay2015community}. Potential changes in climate extremes, because of their heightened societal impacts, are of special concern \citep[e.g.][]{davison1990models, stott2004human, chavez2005generalized, eastoe2009modelling, otto2012reconciling, swain2014extraordinary, singh2014severe, trenberth2015attribution, diffenbaugh2015anthropogenic, huang2015changes, jalbert2017spatiotemporal}. 
However, fully characterizing this evolving natural variability of rare events is intrinsically challenging due to the limited amount of available observations or simulation data.
%identifying changes in climate variability related to human activities is inherently challenging. 
The long equilibration time of the climate system means that on the timescales of interest to human society, the climate state will be evolving, so that its statistical properties are not stationary. Studies of future climate extremes often employ statistical extreme value theory to make inferences about rare events with modest amounts of data \citep{swain2014extraordinary}.

In this work, we study the entire distribution of temperatures in a transient climate, including rare events, by employing quantile regression on an ensemble of simulations of an identical forcing scenario from a single climate model. Given sufficient care in choosing different initial conditions, such a large ensemble will reflect the natural variability of the system, since each simulation will be statistically independent in terms of its natural variability. 
The increased data provided by multiple simulations then admits more confident statements about changes in the statistical behavior of the system than can be made with a single simulation. While the use of ensembles is a relatively recent development, it is growing rapidly \citep[e.g.][]{deser2012uncertainty, deser2012communication,deser2014projecting,fischer2014detection, kay2015community,sriver2015effects, rodgers2015emergence, hagos2016projection}. \citet{deser2012uncertainty}, \citet{deser2012communication} and \citet{fischer2014detection} in particular discuss how ensembles help in distinguishing internal climate variability from anthropogenic effects and allow more accurate estimates for the forced model response.

% XX "Study" was used in two different ways in first two sentences. I changed to say "we will talk about 3 things" - EJM
%One under-exploited advantage of ensembles of multiple simulations is that the increased volume of data allows %XX study of examining the entire distribution of a climate variable.
Ensembles of multiple simulations offer at least three advantages that are currently under-exploited. The most obvious advantage is that the increased data volume allows examining the entire distribution of a climate variable.
 Studies of climate variability to date are generally divided between those that address the center of the distribution 
\citep[e.g.][]{semenov2002secular,raisanen2002co2, kitoh2009changes, screen2014arctic, schneider2015physics}, and those that address its tails \citep[e.g.][]{katz1992extreme, meehl2009relative, northrop2011threshold, davison2012statistical, huser2014space, trenberth2015attribution, huang2015estimating, jalbert2017spatiotemporal}, generally via extreme value theory. %In part the division has arisen from the data scarcity imposed by particular datasets.  %---- END CC
%High-resolution dynamically downscaled model studies \citep{XX,XX,XX} could in principle involve multiple runs, but their computational demands generally limit them to at most a few decades of present and future simulations. 
% XX Those studies 
A more limited body of studies address overall distributional changes in climate variables, but these generally focus on observations or observation-based data products, which are necessarily limited in terms of data amount and therefore require spatial or temporal aggregation \citep{donat2012shifting, stainforth2013mapping, chapman2013estimating, huybers2014us, mckinnon2016changing, rhines2017seasonally}. Aggregating data spatio-temporally requires stationarity assumptions of the signal or explicitly modeling the spatio-temporal dependence. When studying model projections using ensembles, the large amount of data at each location allows us to accurately estimate changes in the distribution of climate variables (e.g. temperature) without spatial aggregation.

A second potential advantage provided by data-rich ensembles is that trends in both means and variability need not be modeled as linear in time \citep{franzke2015local,gao2017quantile}. Typically, analyses assume linear trends,  
but in realistic scenarios, forcings are not linear over centennial timescales, and a linear approximation can be misleading \citep[see for example][]{poppick2017estimating}. The increased data provided by ensembles means that we can consider more flexible statistical models to better represent complex climate responses. %a realistic and often nonlinear scenario. 
As we will show, distributions of daily temperature evolve nonlinearly, and follow different trajectories even as a function of quantiles (i.e. different parts of the distribution). %We therefore need a sound method that can
Analysis methods should therefore be able to take into account nonlinearities both in time and across quantiles.

@@

The ensemble is especially appropriate for the purpose of studying variability because it is based on a ${\sim}10,000$ year pre-industrial control simulation. After a $\sim$4000 year spin-up using constant preindustrial conditions, 50 historical hindcasts (1850-2005) are initialized from snapshots of the coupled model state taken every 100 years, so that the last hindcast is initialized after approximately 9000 years of the control simulation. Each hindcast is then extended to 2100 using the Representative Concentration Pathway (RCP) 8.5 scenario. The 100-year gap between each new initialization ensures nearly independent ensemble members that fully capture internal variability within the coupled system. RCP8.5 corresponds to anthropogenic radiative forcing of roughly 8.5 W m$^{-2}$ by 2100 \citep{moss2010next}. More information about the model and ensemble design can be found in \citet{sriver2015effects}.

@@

% CC -- BEGIN edits below. "Validate" sounds a little more formal than what you're actually doing.
% CC -- if you say " With results for winter, each quantity shows a decent correspondence .." that implies that summer results DONT show a correspondence. 
To evaluate whether the CESM simulations provide sufficiently realistic temperature distributions for the purpose of this analysis, we compare CESM temperatures with those from the ERA-Interim (European Reanalysis) data product \citep{dee2011era}. Figure \ref{model_reanalysis_world} shows the model/reanalysis comparison for winter; for summer see Supplementary Online Material Figure \ref{model_reanalysis_world_JJA}. The model underestimates variability in some places, and produces excessively cold winter temperatures in the Arctic. The resulting exaggerated temperature gradients contribute to excess variability and excessively negative skew in the northern mid-latitudes.
Skewness is proportional to the cube of temperature after subtracting off the average seasonal temperature; see Appendix \ref{app:moments}. %% ejm changed from cubes to cube.
Throughout this work, we will show in-depth analysis from three locations with distinct temperature distributions to highlight our proposed method (\textbf{a}, \textbf{b}, and \textbf{c} shown in Figure \ref{model_reanalysis_world}). See Supplementary Online Material Figure \ref{rea_temp_locations} for comparison of model and reanalysis temperature distributions in both summer and winter for these locations. 
% CC could add a sentence saying this is still OK for your purpose if you want, but I think you can skip that

@@

% CC BEGIN
 As an example of a typical model fit, we show in Figure \ref{Estimate} the seasonal cycle in CESM daily temperatures for three locations, along with estimates of low, median and high quantiles. We show here data from 1850 to demonstrate the seasonal fit rather than that of the long-term trend. All locations show strong seasonal differences in variance that are well-represented by our smooth estimates. Relevant features that are captured include an asymmetrical seasonal cycle in all locations; a clear left skewness in wintertime in all three locations (although most pronounced in the higher-latitude \textbf{a} and \textbf{b}); and a distinct springtime shoulder in the higher-latitude locations. These characteristics show the benefit of explicitly modeling seasonal variations as smoothly varying functions as opposed to a set of four constant functions changing value with season. Nuances like the decrease in winter temperature spread (variability) from early to late winter would not be captured by a piecewise constant model.

@@

% ----- DONE UP THROUGH HERE ---------------- 
\section{Results}
\label{sec:results}

To facilitate comparison to previous studies, we first perform a preliminary analysis where we replicate more standard methods. That is, we examine changes in the aggregate distribution of temperatures over multi-week and multi-month intervals, before we show results from our new approach that calculates responses for individual days. Even the standard analysis readily shows that temperature distributions in the CESM ensemble change over the RCP 8.5 scenario (Figures \ref{location_distributions}, \ref{meanStdSkew_Winter}, and \ref{meanStdSkew_Summer}, which compare the “initial” and “final” time windows 1850-1864 and 2086-2100). Means uniformly shift warmer, but the shapes of the distributions also change in terms of variance and skewness. Figure \ref{location_distributions} shows initial and final distributions in our example locations for aggregated 15-day periods in winter and summer. In at least two of the three depicted locations, it is clear that the distributions are becoming narrower, although quantifying exactly how the tails are changing requires a quantification of the tail size and shape.
%difficult to quantify exactly how the tails are changing.

@@

With a smooth estimate of quantiles of average temperature, we show that the onset of spring, as measured by the first day of the year where the $.5$ quantile estimate reaches $-2.2^{\circ}$C \citep{pearse2017statistical}, occurs earlier in the year as the climate warms in the Detroit area (see Figure \ref{fig:FirstDayAboveFreezing}). The lower quantiles seem to be progessing faster than the $.5$ quantile, with the $.25$ quantile hitting the $-2.2^{\circ}$C mark at a rate of approximately 15 days earlier per decade at present times. Note also that the $.5$ quantile never goes below the threshold after year 2080. It is unclear how to produce the equivalent results using existing methods of segmenting average temperature into seasons. For instance, if we were to look at quantiles of average temperature during winter the edges of the season would pull the overall quantile estimates up and prematurely estimate the onset of spring. Moreover getting information regarding the average temperature of a specific day would be impossible with seasonal averaging.

@@

To clarify the relative contributions of high and low tails to skewness changes, we 
also examine evolving temperature variability in the bulk and tails as a function of seasonality as well as long term change. 
Figure \ref{dq_vs_day} shows absolute variability changes 
for the three example locations \textbf{a}, \textbf{b}, and \textbf{c}
estimated using our quantile model, and for fractional changes see Supplementary Online Material Figure \ref{dq_vs_day_norm}. The uncertainty around our estimates is quantified by resampling the original simulations (with replacement) and recomputing the estimates using this new set of simulations (see Appendix \ref{app:model}\ref{app:uq} for details). In all locations, wintertime skewness changes are driven by the relative changes in IQR and low tails. In the higher-latitude locations \textbf{a} and \textbf{b}, more negative winter skew results because the IQR contracts even more strongly than does the low tail variability. In other words, the low tails ``stick''.  In the lower-latitude location \textbf{c}, more positive winter skew results because the IQR barely changes while the low tail variability contracts strongly.

@@

The complexity of the relationships in Figure \ref{dq_vs_day} also shows how misleading it may be to use a three-month block to represent a season. While all three locations show larger IQR in winter than summer, the transition from winter to summer happens more quickly in some locations than at others, more abruptly in the northernmost location \textbf{a} and more gradually in \textbf{c}. Low-tail variability seasonal transitions are even sharper than those of IQR in \textbf{a} and \textbf{b}, but more gradual in \textbf{c}.
%also has very different profiles between each location. Location \textbf{a} hits peak low-tail variability during the shoulder seasons (spring and fall), while location \textbf{b} shows a more gradual low tail variability rise toward winter. Location \textbf{c} shows less temporal change in the IQR compared with the other locations. Overall, 
In contrast, high-tail variability is more seasonally constant overall than low-tail variability. 
Through these examples, we see how our method offers detailed information about changes in variability across seasons and annual change, usually unavailable when analyzing each season separately.

@@

We present a method to quantify changes in tail variability of temperature with high precision in a transient climate model. Using data from the whole year and the whole span from 1850-2100 we estimate temperature quantiles as a function of seasonality and long term change. Analyzing the whole year simultaneously as opposed to analyzing each season separately allows for more flexible modeling of seasonality. The large ensemble makes it possible to fit such models stably.

@@

The techniques presented in this study are validated in part by the fact that they replicate several prior conclusions made in the literature, e.g. the well-known projected decrease in winter variability in the northern mid-latitudes \citep[e.g.][]{schneider2015physics} most likely due to amplified warming in the arctic \citep{screen2014arctic}. Our approach furthermore allows us to quantify tail variability and give corresponding confidence intervals around our estimates. In the case study of CESM runs analyzed here, we relate the changes in tail variability to changes in skewness of the temperature distributions, and find that in most of the domain analyzed, wintertime skewness changes are driven largely by the relative behavior of IQR and low tails. For example, in much of the continental northern U.S.\ and Canada, the low tail of temperature contracts substantially less than does the overall temperature variability.

@@

The abundance of data available in ensemble simulations relative to single simulations allows using quantile regression to accurately estimate high quantiles, avoiding some of the limitations of extreme value theory. Unlike quantile regression, methods using extreme value theory require making assumptions about the shape of the tail of the distribution. By parameterizing the seasonally time-varying distribution of temperature through smooth functions using the whole year as our domain, we also reveal previously unavailable details  about seasonal transitions. For example, we show here that springtime variability decreases occur later in the year at lower latitudes, and that seasonal transitions in tail variability differ from those in IQR.  While we analyze only temperature here, our method is intended to be general enough to be applied to other climate variables such as precipitation or humidity.
These detailed insights into climate variable distributions may be valuable for risk assessment studies that emphasize extreme events.

>>>>

\affiliation{University of Chicago, Chicago, USA} 
% \affiliation{University of Illinois at Urbana-Champain, Urbana, USA} 
\email{mahaugen@uchicago.edu}
\extraauthor{Ryan L. Sriver}
    \extraaffil{University of Illinois at Urbana-Champaign, Urbana, USA}
% \abstract{
% Understanding future changes in extreme temperature events in a transient climate is inherently challenging. Because a single model simulation is generally insufficient to empirically characterize the statistical properties of underlying processes, many recent efforts have generated ensembles of repeated simulations (with different initial conditions providing independence) to alleviate that data scarcity. We present here a new method for making use of these ensembles that allows characterizing changes in temperature variability for each given time with high confidence by analyzing all temporal observations simultaneously, assuming that temperature distributions vary smoothly both day to day over an annual cycle and over longer secular trends. To demonstrate our method's utility, we analyze an ensemble of 50 simulations of the Community Earth System Model (CESM) under a scenario of increasing radiative forcing to 2100, focusing on North America in particular. The results both verify aspects of climate system behavior known from previous studies, including that wintertime daily temperature variance generally decreases in the continental mid- and high- latitudes ($> 45^{\circ}$), and also elucidate new features. For example, we find that wintertime continental temperature distributions change in a spatial dipole pattern, becoming more left-skewed in the north ($> 45^{\circ}$) and more symmetric at lower latitudes; while changes in summertime distributions show no clear spatial pattern. That is, winter cold extremes become more variable relative to the overall distributions. Although the examples above concern temperature only, the technique is sufficiently general that it can be used to generate precise estimates of distribution changes in a broad range of climate variables by exploiting the power of ensembles.}

\abstract{
Understanding future changes in extreme temperature events in a transient climate is inherently challenging. A single model simulation is generally insufficient to characterize the statistical properties of the underlying physical processes governing the climate. Ensembles of repeated simulations with different initial conditions greatly expand the amount of data available, which in turn allows new approaches for characterizing changes in extremes. Here we present one such new approach, using ensembles that characterize changes in temperature distributions using a continuous representation of seasonality rather than breaking the dataset into seasonal blocks. That is, we assume that temperature distributions evolve smoothly both day-to-day over an annual cycle and year-to-year over longer secular trends. To demonstrate our method's utility, we analyze an ensemble of 50 simulations of the Community Earth System Model (CESM) under a scenario of increasing radiative forcing to 2100, focusing on North America. The results both confirm aspects of climate system behavior known from previous studies and also elucidate new features. Confirming results include that daily temperature bulk variability generally decreases in wintertime in the continental mid- and high-latitudes ($> 40^{\circ}$). One key result is that low wintertime temperatures do not shift as much as the rest of the temperature distribution producing a more negative skew in the overall distribution. Although the examples above concern temperature only, the technique is sufficiently general that it can be used to generate precise estimates of distribution changes in a broad range of climate variables by exploiting the power of ensembles.}

@@

Time series of climate variables have generally been assumed to be separable into two components: randomness inherent in the underlying physical processes, which we call natural variability, and climatic trends, e.g. in the form of forced secular trends that follow from increasing concentrations of greenhouse gases or seasonal trends. Recently, the degree to which natural variability may itself be changing has received significant scientific interest \citep[e.g.][]{trenberth2011attribution, donat2012shifting,deser2012communication, thompson2015quantifying, kay2015community}. Potential changes in climate extremes, because of their heightened societal impacts, are of special concern \citep[e.g.][]{davison1990models, stott2004human, chavez2005generalized, eastoe2009modelling, otto2012reconciling, swain2014extraordinary, singh2014severe, trenberth2015attribution, diffenbaugh2015anthropogenic, huang2015changes, jalbert2017spatiotemporal}. 
However, fully characterizing this evolving natural variability of rare events is intrinsically challenging due to the limited amount of available observations or simulation data. The long equilibration time of the climate system means that on the timescales of interest to human society, the climate state will be evolving, so that its statistical properties are not stationary. Studies of future climate extremes often employ statistical extreme value theory to make inferences about rare events with modest amounts of data \citep{swain2014extraordinary}.

In this work, we study the entire distribution of temperatures in a transient climate, including rare events, by employing quantile regression on an ensemble of simulations of an identical forcing scenario from a single climate model. Sufficient sampling of the initial conditions' uncertainty will reflect the natural variability of the climate system, since each simulation is statistically independent in terms of its natural variability. 
The increased data provided by multiple simulations enables more confident statements about changes in the statistical behavior of the system than can be made with a single simulation. The use of initial conditions for characterizing internal variability is growing rapidly \citep[e.g.][]{deser2012uncertainty, deser2012communication,deser2014projecting,fischer2014detection, kay2015community,sriver2015effects, rodgers2015emergence, hagos2016projection}. \citet{deser2012uncertainty}, \citet{deser2012communication} and \citet{fischer2014detection} in particular discuss how ensembles help distinguish internal climate variability from anthropogenic effects on temperature changes and allow more comprehensive estimates of the model's temperature response to radiative forcing.

% XX "Study" was used in two different ways in first two sentences. I changed to say "we will talk about 3 things" - EJM
%One under-exploited advantage of ensembles of multiple simulations is that the increased volume of data allows %XX study of examining the entire distribution of a climate variable.
Large single model ensembles offer at least three advantages over a single simulation of a climate model. The most obvious advantage is that the increased data volume allows examining the entire distribution of a climate variable.
 Studies of climate variability to date are generally divided between those that address the center of the distribution 
\citep[e.g.][]{semenov2002secular,raisanen2002co2, kitoh2009changes, screen2014arctic, schneider2015physics}, and those that address its tails \citep[e.g.][]{katz1992extreme, meehl2009relative, northrop2011threshold, davison2012statistical, huser2014space, trenberth2015attribution, huang2015estimating, jalbert2017spatiotemporal}, generally via extreme value theory. %In part the division has arisen from the data scarcity imposed by particular datasets.  %---- END CC
%High-resolution dynamically downscaled model studies \citep{XX,XX,XX} could in principle involve multiple runs, but their computational demands generally limit them to at most a few decades of present and future simulations. 
% XX Those studies 
A more limited body of studies address overall distributional changes in climate variables, but these generally focus on observations or observation-based data products, which are necessarily limited in terms of data amount and therefore require spatial or temporal aggregation \citep{donat2012shifting, stainforth2013mapping, chapman2013estimating, huybers2014us, mckinnon2016changing, rhines2017seasonally}. Aggregating data spatio-temporally requires stationarity assumptions of the signal or explicitly modeling the spatio-temporal dependence. When studying model projections using ensembles, the large amounts of data at each location allows us to estimate changes in the distribution of climate variables (e.g. temperature) without spatial aggregation.

A second advantage provided by large single model ensembles is that trends in both means and variability need not be modeled as linear in time \citep{franzke2015local,gao2017quantile}. Forcings are not linear over centennial timescales, and a linear approximation can be misleading \citep[see for example][]{poppick2017estimating}. The increased data provided by ensembles means that we can consider more flexible statistical models to represent complex climate responses. %a realistic and often nonlinear scenario. 
As we will show, distributions of daily temperature evolve nonlinearly, and follow different trajectories even as a function of quantiles (i.e. different parts of the distribution). %We therefore need a sound method that can
Analysis methods should therefore be able to take into account nonlinearities both in time and across quantiles.

@@

The ensemble is appropriate for the purpose of studying coupled internal climate variability because it is based on a ${\sim}10,000$ year pre-industrial control simulation. After a $\sim$4000 year spin-up using constant preindustrial conditions, 50 historical hindcasts (1850-2005) are initialized from snapshots of the coupled model state taken every 100 years, so that the last hindcast is initialized after approximately 9000 years of the control simulation. Each hindcast is then extended to 2100 using the Representative Concentration Pathway (RCP) 8.5 scenario. The 100-year gap between each new initialization ensures nearly independent ensemble members that fully capture internal variability within the coupled system. RCP8.5 corresponds to anthropogenic radiative forcing of roughly 8.5 W m$^{-2}$ by 2100 \citep{moss2010next}. More information about the model and ensemble design can be found in \citet{sriver2015effects}.

@@

% CC -- BEGIN edits below. "Validate" sounds a little more formal than what you're actually doing.
% CC -- if you say " With results for winter, each quantity shows a decent correspondence .." that implies that summer results DONT show a correspondence. 
To evaluate whether the CESM simulations provide sufficiently realistic temperature distributions for the current analysis, we compare CESM temperatures with those from the ERA-Interim (European Reanalysis) data product \citep{dee2011era}. Figure \ref{model_reanalysis_world} shows the model/reanalysis comparison for winter; for summer see Supplementary Online Material Figure \ref{model_reanalysis_world_JJA}. The model underestimates variability in some places, and produces excessively cold winter temperatures in the Arctic. The resulting temperature gradients contribute to excess variability and negative skew in the northern mid-latitudes.
Skewness is proportional to the cube of temperature after subtracting off the average seasonal temperature; see Appendix \ref{app:moments}. %% ejm changed from cubes to cube.
Throughout this work, we will show in-depth analysis from three locations with distinct temperature distributions to highlight our proposed method (\textbf{a}, \textbf{b}, and \textbf{c} shown in Figure \ref{model_reanalysis_world}). See Supplementary Online Material Figure \ref{rea_temp_locations} for comparison of model and reanalysis temperature distributions in both summer and winter for these locations. 
% CC could add a sentence saying this is still OK for your purpose if you want, but I think you can skip that

@@

% CC BEGIN
 As an example of a typical model fit, we show in Figure \ref{Estimate} the seasonal cycle in CESM daily temperatures for three locations, along with estimates of low, median and high quantiles. We show here data from 1850 to demonstrate the seasonal fit rather than that of the long-term trend. All locations show strong seasonal differences in variance that are well-represented by our smooth estimates. Relevant features that are captured include an asymmetrical seasonal cycle in all locations; a clear left skewness in wintertime in all three locations (although most pronounced in the higher-latitude \textbf{a} and \textbf{b}); and a distinct springtime shoulder in the higher-latitude locations. These characteristics show the benefit of explicitly modeling seasonal variations as smoothly varying functions as opposed to a set of four constant functions changing with the seasons. Nuances like the decrease in winter temperature spread (variability) from early to late winter would not be captured by a piecewise constant model.

@@

% ----- DONE UP THROUGH HERE ---------------- 
\section{Results and Discussion}
\label{sec:results}

To facilitate comparison with previous studies, we first perform a preliminary analysis where we replicate more standard methods. That is, we examine changes in the aggregate distribution of temperatures over multi-week and multi-month intervals, before we show results from our new approach that calculates responses for individual days. Even the standard analysis readily shows that temperature distributions in the CESM ensemble change over the RCP 8.5 scenario (Figures \ref{location_distributions}, \ref{meanStdSkew_Winter}, and \ref{meanStdSkew_Summer}, which compare the “initial” and “final” time windows 1850-1864 and 2086-2100). Means uniformly shift to warmer temperatures, but the shapes of the distributions also change in terms of variance and skewness. Figure \ref{location_distributions} shows initial and final distributions in our example locations for aggregated 15-day periods in winter and summer. In at least two of the three depicted locations, it is clear that the distributions are becoming narrower, although quantifying exactly how the tails are changing requires a quantification of the tail size and shape.
%difficult to quantify exactly how the tails are changing.

@@

With a smooth estimate of quantiles of average temperature, we show that the onset of spring, as measured by the first day of the year where the $.5$ quantile estimate reaches $-2.2^{\circ}$C \citep{pearse2017statistical}, occurs earlier in the year as the climate warms in the Detroit area (see Figure \ref{fig:FirstDayAboveFreezing} where location \textbf{b} is analyzed). The lower quantiles progress faster than the $.5$ quantile, with the $.25$ quantile hitting the $-2.2^{\circ}$C mark at a rate of approximately 15 days earlier per decade at present times. Note also that the $.5$ quantile never goes below the threshold after year 2080. It is unclear how to produce the equivalent results using existing methods of segmenting average temperature into seasons. For instance, if we were to look at quantiles of average temperature during winter the edges of the season would pull the overall quantile estimates up and prematurely estimate the onset of spring. Moreover getting information regarding the average temperature of a specific day would be impossible with seasonal averaging.

@@

To clarify the relative contributions of high and low tails to skewness changes, we 
also examine evolving temperature variability in the bulk and tails as a function of seasonality as well as long term change. 
Figure \ref{dq_vs_day} shows absolute variability changes 
for the three example locations \textbf{a}, \textbf{b}, and \textbf{c}
estimated using our quantile model, and for fractional changes see Supplementary Online Material Figure \ref{dq_vs_day_norm}. The uncertainty around our estimates is quantified by resampling the original simulations (with replacement) and recomputing the estimates using this new set of simulations (see Appendix \ref{app:model}\ref{app:uq} for details). In all locations, wintertime skewness changes are driven by the relative changes in IQR and low tails. In the higher-latitude locations \textbf{a} and \textbf{b}, more negative winter skew results because the IQR contracts even more strongly than does the low tail variability. In other words, the low tails ``stick''.  In the lower-latitude location \textbf{c}, more positive winter skew results because the IQR changes slightly while the low tail variability contracts strongly.

@@

The complexity of the relationships in Figure \ref{dq_vs_day} also shows how misleading it may be to use a three-month block to represent a season. While all three locations show larger IQR in winter than summer, where the transition from winter to summer happens more quickly at some locations than at others. This transition takes place abruptly in the northernmost location \textbf{a} and more gradually in \textbf{c}. Low-tail variability seasonal transitions are even sharper than those of IQR in \textbf{a} and \textbf{b}, but more gradual in \textbf{c}.
%also has very different profiles between each location. Location \textbf{a} hits peak low-tail variability during the shoulder seasons (spring and fall), while location \textbf{b} shows a more gradual low tail variability rise toward winter. Location \textbf{c} shows less temporal change in the IQR compared with the other locations. Overall, 
In contrast, high-tail variability is more seasonally constant overall than low-tail variability. 
Through these examples, we see how our method offers detailed information about changes in variability across seasons and annual change, usually unavailable when analyzing each season separately.

@@

We present a method to quantify changes in tail variability of temperature with high precision using a 50-member single climate model ensemble (\citep{sriver2015effects, hogan2017analyzing, vega2017analysis}. Using data from the whole year and the whole span from 1850-2100 we estimate temperature quantiles as a function of seasonality and long term change. Analyzing the whole year simultaneously as opposed to analyzing each season separately allows for more flexible modeling of seasonality. Fitting these models stably can be achieved using large model ensembles sampling initial conditions uncertainty (i.e. internal variability).

@@

The techniques presented in this study are supported in part because they replicate several prior conclusions made in the literature, e.g. the well-known projected decrease in winter variability in the northern mid-latitudes \citep[e.g.][]{schneider2015physics} most likely due to amplified warming in the arctic \citep{screen2014arctic}. Our approach furthermore enables quantification of tail variability and corresponding confidence intervals. In the case study of CESM runs analyzed here, we relate the changes in tail variability to changes in skewness of the temperature distributions, and we find that in most of the domain analyzed, wintertime skewness changes are driven largely by the relative behavior of IQR and low tails. For example, in much of the continental northern U.S.\ and Canada, the low tail of temperature contracts substantially less than does the overall temperature variability.

@@

The abundance of data available in large single model ensembles relative to single simulations allows using quantile regression to accurately estimate high quantiles within a single model structure, avoiding some of the limitations of extreme value theory. Unlike quantile regression, methods using extreme value theory require making assumptions about the shape of the tail of the distribution. By parameterizing the seasonally time-varying distribution of temperature through smooth functions using the whole year as our domain, we also reveal previously unavailable details  about seasonal transitions. For example, we show here that springtime variability decreases occur later in the year at lower latitudes, and that seasonal transitions in tail variability differ from those in IQR.  While we analyze only temperature here, our method is intended to be general enough to be applied to other climate variables such as precipitation or humidity.
These detailed insights into climate variable distributions may be valuable for risk assessment studies that emphasize extreme events.
